\documentclass[12pt,english,a4paper]{article}

\usepackage[utf8]{inputenc}          % Allows UTF-8 encoded characters in the .tex-file.
\usepackage{babel,csquotes,textcomp} % Set LaTeX to structure the content following international academic standards.

% Include the TikZ-package for drawing figures.
%\usepackage{tikz}
%\usetikzlibrary{shapes.geometric,arrows}
%\usepackage{courier}

% Creates clickable hyperlinks in the PDF-document.
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{pdfpages}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{color}

\usepackage[
    backend=biber,
    style=alphabetic,
    sorting=ynt
]{biblatex}
\addbibresource{refs.bib}
% Removes the dots in the table of contents.
%\usepackage[titles]{tocloft} 
%\renewcommand{\cftdot}{}

% Enables bibtex for keeping track of the document's reference list.
%\usepackage[style=numeric-comp,backend=biber,sortcites]{biblatex} 
\urlstyle{sf}
%\addbibresource{bibliography.bib} % The text-file that contains the document's references.

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ %
  basicstyle=\ttfamily\small,     
  backgroundcolor=\color{white},   % choose the background color
  breaklines=true,                 % automatic line breaking only at whitespace
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  keywordstyle=\color{blue},       % keyword style
  stringstyle=\color{mymauve},     % string literal style
}
%\nocite{*}	% This should be removed or the bibliography needs to be updated. 

%opening
\title{GPS Spoofing and the importance of correct time}
\author{Aril Johannes Schultzen}

\begin{document}
\maketitle
\thispagestyle{empty}
\setcounter{page}{0}
\newpage
\tableofcontents
\thispagestyle{empty}
\setcounter{page}{0}
\newpage
\thispagestyle{empty}
\setcounter{page}{0}

\begin{abstract}
Currently, this is not much more but comments and notes in order for me to write an essay, an essay meant for me to prepare for my work with my master thesis. This document does not represent the quality of the finished product in any way. Example cite \cite{KandR}
\end{abstract}

\newpage
\clearpage
\setcounter{page}{1}

\section{Global Positioning System: A short introduction}
Most of use have at some time in some way used GPS....go on. Elaborate.

\section{Phasor Measurement Units}

\section{Signal interference: Jamming}
\subsection{Locking on}
\subsection{Jamming: How to?}

By emitting a high-power signal at the GPS frequency, one can interfere with the signals emitted by the GPS satellites, effectively denying GPS receivers use of these signals. These signals are already weak considering their travel from space. Such an "attack", although effective, is pretty naive and easily recognized by the jammed party. After all, if you don't have a signal, you are probably being jammed. A more sophisticated approach is to emit a counterfeit signal in order to manipulate the receivers reported time or even position. IRAN, DRONE SPOOF STORY!. 

\subsection{Detecting Jamming}

\subsection{Preventing Jamming}
\subsubsection{Encryption}

\section{Notes}
A way of detecting a clumsy attempt at jamming, would be to enter a "Jam-mode" as soon as the location changes. The PMUs are not supposed to be moved. Why would they anyway? Moving power-plants? Discuss with Harald.

\newpage
\printbibliography[title={Complete Bibliography},heading=bibintoc]

\end{document}